Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv('H:/Udacity/Nanodegree Data Analyst/Exploratory Data Analysis - R/Skript - R Basic/pseudo_facebook.tsv', sep = '\t')
names(pf)
str(pf)

# Create a scatter plot with two continoues variables
qplot(x = age, y = friend_count, data = pf)

# Similar result, without x/y --> qplot knows what variables should be on x or y
qplot(age, friend_count, data = pf)

```

***

#### What are some things that you notice right away?
Response:
Every group of age has in the most cases 500 up to 1000 friends. The most friends have young people between 15 and 25. But there are some peaks at over 60 years and around the age of 100.
***

### ggplot Syntax
Notes:
ggplot is better by more complecated plots than qplot

```{r ggplot Syntax}
# Generate the similar Plot like qplot
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point() + 
  xlim(13, 90)

summary(pf$age)
# Min.    1st Qu. Median  Mean    3rd Qu. Max. 
# 13.00   20.00   28.00   37.28   50.00   113.00 
```

***

### Overplotting
Notes: 
Overplotting means that is difficult to analyze the data because there are too many points in a region.
alpha = 1 / 20 means that 1 point stays for 20 points in the same region
```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_jitter(alpha = 1/20) + 
  xlim(13, 90)
```

#### What do you notice in the plot?
Response:
That the count of friends increase till 22 and than decrease.
***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_jitter(alpha = 1/20) + 
  xlim(13, 90) + 
  coord_trans(x = "sqrt")
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_jitter(alpha = 1/20, position = position_jitter(h = 0)) + 
  xlim(13, 90) + 
  coord_trans(x = 'sqrt')
```

#### What do you notice?
the y axes has a wider range from 0-30 and 30-50 then the other ones

position = position_jitter(h = 0) makes a break at zero to see negative friend_counts
***

### Alpha and Jitter
Notes:
# This programming assignment
# will not be graded, but when you
# submit your code, the assignment
# will be marked as correct. By submitting
# your code, we can add to the feedback
# messages and address common mistakes
# in the Instructor Notes.

# You can assess your work by watching
# the solution video.


# Examine the relationship between
# friendships_initiated (y) and age (x)
# using the ggplot syntax.

# We recommend creating a basic scatter
# plot first to see what the distribution looks like.
# and then adjusting it by adding one layer at a time.

# What are your observations about your final plot?

# Remember to make adjustments to the breaks
# of the x-axis and to use apply alpha and jitter.

```{r Alpha and Jitter}
ggplot(aes(x = age, y = friendships_initiated), data = pf) + 
  geom_jitter(alpha = 1/10, position = position_jitter(h = 0)) +
  xlim(13, 90) + 
  coord_trans(y = 'sqrt')
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:
Functions that you normally use:
filter()
group_by()
mutate()
arrange()

n() gives the number of the searched dataset, this function can only be used inside the summarise function

```{r Conditional Means}
#install.packages("dplyr")
#library(dplyr)

age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups, 
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n())

head(pf.fc_by_age)

pf.fc_by_age <- arrange(pf.fc_by_age, age)
head(pf.fc_by_age, 20)

```

### Conditional Means Alternate Code 
%>% chain function on a data set
```{r Conditional Means Plot}
pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n()) %>%
  arrange(pf.fc_by_age, age)

head(pf.fc_by_age, 20)
```

***

Create your plot!
# This programming assignment
# will not be graded, but when you
# submit your code, the assignment
# will be marked as correct. By submitting
# your code, we can add to the feedback
# messages and address common mistakes
# in the Instructor Notes.

# You can assess your work by watching
# the solution video.


# Plot mean friend count vs. age using a line graph.
# Be sure you use the correct variable names
# and the correct data frame. You should be working
# with the new data frame created from the dplyr
# functions. The data frame is called 'pf.fc_by_age'.

# Use geom_line() rather than geom_point to create
# the plot. You can look up the documentation for
# geom_line() to see what it does.
```{r Conditional Means Plot}
ggplot(aes(age, friend_count_mean), data = pf.fc_by_age) + 
  geom_line() +
  xlim(13, 100)
```

***

### Overlaying Summaries with Raw Data
Notes:
Add line from Conditional Means
```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_jitter(alpha = 0.05, position = position_jitter(h = 0),
              color = 'orange') + 
  #xlim(13, 90) + 
  coord_cartesian(xlim = c(13, 90)) +
  coord_trans(y = 'sqrt') +
  geom_line(stat = 'summary', fun.y = mean) + 
  geom_line(stat = 'summary', fun.y = quantile, fun.args=list(probs=0.1),
            linetype = 2, color = 'blue') + 
  geom_line(stat = 'summary', fun.y = quantile, fun.args=list(probs=0.5),
            color = 'blue') + 
  geom_line(stat = 'summary', fun.y = quantile, fun.args=list(probs=0.9),
            linetype = 2, color = 'blue')


```

#### What are some of your observations of the plot?
Response:
The distribution and quartiles changed.
***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age, pf$friend_count, method = 'pearson')

with(pf, cor.test(age, friend_count, method = 'pearson'))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:
-0.027
***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:
with(subset(pf, age <= 70), cor.test(age, friend_count, method = 'pearson'))  # Result: -0.1717245
with(subset(pf, age <= 70), cor.test(age, friend_count, method = 'spearman')) # Result: -0.2552934
***

## Create Scatterplots
Notes:
Compare www_likes_received with likes_received
likes_received shold be higher than www_likes_received
# This programming assignment
# will not be graded, but when you
# submit your code, the assignment
# will be marked as correct. By submitting
# your code, we can add to the feedback
# messages and address common mistakes
# in the Instructor Notes.

# You can assess your work by watching
# the solution video.


# Create a scatterplot of likes_received (y)
# vs. www_likes_received (x). Use any of the
# techniques that you've learned so far to
# modify the plot.
```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
  geom_point()
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
  geom_point() + 
  xlim(0, quantile(pf$www_likes_received, 0.95)) + 
  ylim(0, quantile(pf$likes_received, 0.95)) + 
  geom_smooth(method = 'lm', color = 'red')

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
with(pf, cor.test(www_likes_received, likes_received, method = 'pearson'))
```

Response:
0.948
***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
data(Mitchell)
?Mitchell
```

Create your plot!
# This programming assignment
# will not be graded, but when you
# submit your code, the assignment
# will be marked as correct. By submitting
# your code, we can add to the feedback
# messages and address common mistakes
# in the Instructor Notes.

# You can assess your work by watching
# the solution video.


# Create a scatterplot of temperature (Temp)
# vs. months (Month).
```{r Temp vs Month}
ggplot(aes(x = Temp, y = Month), data = Mitchell) + 
  geom_point() #+ 
  #xlim(0, quantile(pf$www_likes_received, 0.95)) + 
  #ylim(0, quantile(pf$likes_received, 0.95)) + 
  #geom_smooth(method = 'lm', color = 'red')

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.
0 (-2 to 2)

b. What is the actual correlation of the two variables?
(Round to the thousandths place)
0.05747063

```{r Noisy Scatterplots}
with(Mitchell, cor.test(Temp, Month))
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(x = Month, y = Temp), data = Mitchell) + 
  geom_point()+
  scale_x_continuous(breaks = seq(0, 203, 12))
```

***

### A New Perspective

What do you notice?
Response:
That there are season in Alaska, the temperature goes in between 12 months up and down.

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
pf$age_with_months <- with(pf, age + (1 - dob_month / 12))
head(pf$age_with_months)


```

***

### Age with Months Means
# Create a new data frame called
# pf.fc_by_age_months that contains
# the mean friend count, the median friend
# count, and the number of users in each
# group of age_with_months. The rows of the
# data framed should be arranged in increasing
# order by the age_with_months variable.

# For example, the first two rows of the resulting
# data frame would look something like...

# age_with_months  friend_count_mean	friend_count_median	n
#              13            275.0000                   275 2
#        13.25000            133.2000                   101 11


# See the Instructor Notes for two hints if you get stuck.
# This programming assignment will automatically be graded.
```{r Age with Months Means}

```

Programming Assignment
```{r Programming Assignment}
pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n()) #%.%
  #arrange(pf$age_with_months)

head(pf.fc_by_age_months, 20)

```

***

### Noise in Conditional Means
# This programming assignment will not be graded, but when you submit your code,
# the assignment will be marked as correct. By submitting your code, we can add
# to the feedback messages and address common mistakes in the Instructor Notes.

# You can assess your work by watching the solution video.


# Create a new line plot showing friend_count_mean versus the new variable,
# age_with_months. Be sure to use the correct data frame (the one you created
# in the last exercise) AND subset the data to investigate users with ages less
# than 71.
```{r Noise in Conditional Means}
ggplot(aes(x = age_with_months, y = friend_count_mean), 
       data = subset(pf.fc_by_age_months, age_with_months < 71))+ 
  geom_line()
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(x = age, y = friend_count_mean), 
             data = subset(pf.fc_by_age, age < 71)) + 
  geom_line() +
  geom_smooth() 

p2 <- ggplot(aes(x = age_with_months, y = friend_count_mean), 
       data = subset(pf.fc_by_age_months, age_with_months < 71))+ 
  geom_line() +
  geom_smooth()
  

p3 <- ggplot(aes(x = round(age/5)*5, y = friend_count), 
       data = subset(pf, age < 71))+ 
  geom_line(stat = 'summary', fun.y = mean)

library(gridExtra)
grid.arrange(p2, p1, p3, ncol = 1)

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

