pf <- read.csv('H:/Udacity/Nanodegree Data Analyst/Exploratory Data Analysis - R/Skript - R Basic/pseudo_facebook.tsv', sep = '\t')

names(pf)
# Output: 
#[1] "userid"                "age"                   "dob_day"               "dob_year"              "dob_month"            
#[6] "gender"                "tenure"                "friend_count"          "friendships_initiated" "likes"                
#[11] "likes_received"        "mobile_likes"          "mobile_likes_received" "www_likes"             "www_likes_received"   



Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
ggplot(aes(x = gender, y = age), data = subset(pf, !is.na(gender))) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

ggplot(aes(x = age, y = friend_count), data = subset(pf, !is.na(gender))) + 
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)

```

# Write code to create a new data frame,
# called 'pf.fc_by_age_gender', that contains
# information on each age AND gender group.

# The data frame should contain the following variables:

#    mean_friend_count,
#    median_friend_count,
#    n (the number of users in each age and gender grouping)

# Here is an example of the structure of your data frame. Your
# data values will be different. Note that if you are grouping by
# more than one variable, you will probably need to call the
# ungroup() function. 

#   age gender mean_friend_count median_friend_count    n
# 1  13 female          247.2953                 150  207
# 2  13   male          184.2342                  61  265
# 3  14 female          329.1938                 245  834
# 4  14   male          157.1204                  88 1201
Notes:
```{r Plotting Conidtional Quiz}
suppressMessages(library(dplyr))

pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
          n = n()) %>%
  ungroup() %>%
  arrange(age)

head(pf.fc_by_age_gender)
```

***
### Plotting Conidtional Quiz
### Plotting Conditional Summaries

# Create a line graph showing the
# median friend count over the ages
# for each gender. Be sure to use
# the data frame you just created,
# pf.fc_by_age_gender.


```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = median_friend_count), data = pf.fc_by_age_gender) + 
  geom_line(aes(color = gender))

```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
#install.packages('reshape2')
#library(reshape2)

#create a new variable
# d = dataframe
# cast = for array or matrix
# specify the dataframe which you will change
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, 
                                  age ~ gender,
                                  value.var = 'median_friend_count')

head(pf.fc_by_age_gender.wide)

```


***

### Ratio Plot
# Plot the ratio of the female to male median
# friend counts using the data frame
# pf.fc_by_age_gender.wide.

# Think about what geom you should use.
# Add a horizontal line to the plot with
# a y intercept of 1, which will be the
# base line. Look up the documentation
# for geom_hline to do that. Use the parameter
# linetype in geom_hline to make the
# line dashed.

# The linetype parameter can take the values 0-6:
# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash
Notes:

```{r Ratio Plot}
ggplot(data = pf.fc_by_age_gender.wide, aes(x = age, y = female/male)) + 
  geom_line() + 
  geom_hline(yintercept = 1, alpha = 0.3, linetype = 2)


```

***

### Third Quantitative Variable
# Create a variable called year_joined
# in the pf data frame using the variable
# tenure and 2014 as the reference year.

# The variable year joined should contain the year
# that a user joined facebook.

# See the Instructor Notes for three hints if you get
# stuck. Scroll down slowly to see one hint at a time
# if you would like some guidance.


```{r Third Quantitative Variable}
# create a variable year_joined and assign it to pf
# floor gives the value without decimals
pf$year_joined <- floor(2014 - pf$tenure/365)
head(pf$year_joined)


# pf has now one more variable
```

***

### Cut a Variable
# Create a new variable in the data frame
# called year_joined.bucket by using
# the cut function on the variable year_joined.

# You need to create the following buckets for the
# new variable, year_joined.bucket

#        (2004, 2009]
#        (2009, 2011]
#        (2011, 2012]
#        (2012, 2014]

# Note that a parenthesis means exclude the year and a
# bracket means include the year.

# Look up the documentation for cut or try the link
# in the Instructor Notes to accomplish this task.
Notes:

```{r Cut a Variable}
summary(pf$year_joined)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
#   2005    2012    2012    2012    2013    2014       2 


table(pf$year_joined)
# 2005  2006  2007  2008  2009  2010  2011  2012  2013  2014 
#    9    15   581  1507  4557  5448  9860 33366 43588    70 

#create new variable bucket
pf$year_joined.bucket <- cut(pf$year_joined, c(2004, 2009, 2011, 2012, 2014))
table(pf$year_joined.bucket)
#(2004,2009] (2009,2011] (2011,2012] (2012,2014] 
#       6669       15308       33366       43658 

```

***

### Plotting it All Together
# Create a line graph of friend_count vs. age
# so that each year_joined.bucket is a line
# tracking the median user friend_count across
# age. This means you should have four different
# lines on your plot.

# You should subset the data to exclude the users
# whose year_joined.bucket is NA.

# should look similar to the following code
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)

Notes:

```{r Plotting it All Together}
#table(pf$year_joined.bucket, useNA = 'ifany')
#(2004,2009] (2009,2011] (2011,2012] (2012,2014]        <NA> 
#       6669       15308       33366       43658           2 

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median) 

```

***

### Plot the Grand Mean
# Write code to do the following:

# (1) Add another geom_line to code below
# to plot the grand mean of the friend count vs age.

# (2) Exclude any users whose year_joined.bucket is NA.

# (3) Use a different line type for the grand mean.

# As a reminder, the parameter linetype can take the values 0-6:

# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash


# The code from the last programming exercise should
# be your starter code!
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = mean ,linetype = 6)
```

***

### Friending Rate
Notes:
What is the median friend rate?
0.2205

What is the max friend rate?
417

```{r Friending Rate}
with(subset(pf, tenure >= 1), summary(friend_count / tenure))

```

***

### Friendships Initiated
# Create a line graph of mean of friendships_initiated per day (of tenure)
# vs. tenure colored by year_joined.bucket.

# You need to make use of the variables tenure,
# friendships_initiated, and year_joined.bucket.

# You also need to subset the data to only consider user with at least
# one day of tenure.

Notes:

What is the median friend rate?
0.025
What is the maximum friend rate?
417
```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)


# Instead of geom_line(), use geom_smooth() to add a smoother to the plot.
# You can use the defaults for geom_smooth() but do color the line
# by year_joined.bucket

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket))
  
```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yogurt <- read.csv(file = 'H:/Udacity/Nanodegree Data Analyst/Exploratory Data Analysis - R/Skript - R Basic/yogurt.csv', sep=';', dec = '.')
str(yogurt)


yogurt$id <- factor(yogurt$id)
str(yogurt)

# Quiz:
ggplot(data = yogurt, aes(x = price)) + 
  geom_histogram(fill = '#00f5ff')


```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
summary(yogurt)

length(unique(yogurt$price))

table(yogurt$price)

#Quiz:
# Create a new variable called all.purchases,
# which gives the total counts of yogurt for
# each observation or household.

# One way to do this is using the transform
# function. You can look up the function transform
# and run the examples of code at the bottom of the
# documentation to figure out what it does.

# The transform function produces a data frame
# so if you use it then save the result to 'yo'!

# OR you can figure out another way to create the
# variable.

yogurt$all.purchases 
yogurt <- transform(yogurt, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)

head(yogurt)

```

***

### Prices over Time
# Create a scatterplot of price vs time.

# This will be an example of a time series plot.

# Resolve overplotting issues by using
# techniques you learned in Lesson 4.

# What are some things that you notice?

```{r Prices over Time}
ggplot(aes(x = time, y = price), data = yogurt) + 
  geom_jitter(alpha = 1/4, shape = 21, fill = '#00f5ff')

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
# set the seeds of reproducible results
set.seed(4230)
sample.ids <- sample(levels(yogurt$id), 16)


ggplot(aes(x = time, y = price), 
       data = subset(yogurt, id %in% sample.ids)) + 
  facet_wrap( ~ id) + 
  geom_line() + 
  geom_point(aes(size = all.purchases), pch = 1)


#Quiz:
# set the seeds of reproducible results
set.seed(1000)
sample.ids <- sample(levels(yogurt$id), 16)


ggplot(aes(x = time, y = price), 
       data = subset(yogurt, id %in% sample.ids)) + 
  facet_wrap( ~ id) + 
  geom_line() + 
  geom_point(aes(size = all.purchases), pch = 1)

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

```{r}
library(GGally)
theme_set(theme_minimal(20))

# set the seeds of reproducible results
set.seed(1836)
pf_subset() <- pf[, c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000),])

#generate the scatterplot matrix takes more than an hour

```
***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("H:/Udacity/Nanodegree Data Analyst/Exploratory Data Analysis - R/Skript - R Basic/nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

